\texttt{nuance}: Detection of planetary transits in the presence of correlated noise

### 2023-01-23 simple/latex/ms.tex
add proof and test

% Define document class
\documentclass{aastex631}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}

@@

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
        \caption{\TODO}
        \label{fig:snr_detrend}
    \end{centering}
\end{figure}

@@

i.e., given a depth $D$, the likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$. Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e. transits with different depths), we decompose
\begin{equation}\label{eq:non_part_of_per}
    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta
\end{equation}
where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$. In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
\\\\
Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood:
$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k G_k(\tilde\Delta)$$
where $G_k$ is the probability density function of $\mathcal{N}(\Delta_k, \sigma_k^2)$. 
\\\\
As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2) \quad\text{i.e.}\quad \Delta \sim \mathcal{N}(\Delta, \sigma^2) $$
a normal distribution defined by
\begin{equation}\label{eq:params}
\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad
\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
\end{equation}
$$p(\tilde\Delta | \Delta) =  \left(\prod_{\sumTk}^K \mathcal{L}_i \right) G(\tilde\Delta)$$
where $G$ is the probability density function of $\mathcal{N}(\Delta, \sigma^2)$. We can now rewrite \autoref{eq:non_part_of_per} as
\begin{equation}\label{eq:ice_int}
    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i \right)  \int G_k(\tilde\Delta)\, G(\tilde\Delta)\, d\tilde\Delta
\end{equation}
The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
$$p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}$$
And \autoref{eq:goal} gives 
\begin{equation}\label{eq:result}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
\end{equation}
the log-likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$. In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.

>>>>

% Define document class
\documentclass{aastex631}
% \usepackage{showyourwork}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}

@@

% Cortex style
\input{\cortex}

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[height=9cm]{../../workflows/cleaning_snr/figures/simu1/result.pdf}
        \caption{\TODO{}, and redo plot with top arrows and fix $tau_v$ $delta_v$}
        \label{fig:snr_detrend}
    \end{centering}
\end{figure}

@@

i.e., given a depth $D$, the likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$. Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e. transits with different depths), we decompose
\begin{equation}\label{eq:non_part_of_per}
    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta
\end{equation}
where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$. In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
\\\\
Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood:
$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k G_k(\tilde\Delta)$$
where $G_k$ is the probability density function of $\mathcal{N}(\Delta_k, \sigma_k^2)$. 
\\\\
As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2) \quad\text{i.e.}\quad \Delta \sim \mathcal{N}(\Delta, \sigma^2) $$
a normal distribution defined by
\begin{equation}\label{eq:params}
\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad
\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
\end{equation}
$$p(\tilde\Delta | \Delta) =  \left(\prod_{\sumTk}^K \mathcal{L}_i \right) G(\tilde\Delta)$$
where $G$ is the probability density function of $\mathcal{N}(\Delta, \sigma^2)$. We can now rewrite \autoref{eq:non_part_of_per} as
\begin{equation}\label{eq:ice_int}
    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i \right)  \int G_k(\tilde\Delta)\, G(\tilde\Delta)\, d\tilde\Delta
\end{equation}
The product of two normal distributions is also a normal distribution, so that the integral in \autoref{eq:ice_int} can be obtained analytically. We have
\begin{proof}{integral_sympy}
    p(\bm{f} \vert T_k, D, \Delta) =   \mathcal{L}_k \left(\prod_{\sumTk}^K \mathcal{L}_i  \right) \frac{1}{\sqrt{2\pi \left(\sigma^{2} + \sigma_{k}^{2}\right)}} e^{- \frac{1}{2}\frac{\left(\Delta -  \Delta_{k}\right)^{2}}{\sigma^{2} + \sigma_{k}^{2}}}
\end{proof}
And \autoref{eq:goal} gives 
\begin{equation}\label{eq:result}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{K}{2}\ln(2\pi) - \frac{1}{2}\sum_k^K \left( \frac{\left(\Delta_{k} -
    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}} + \ln{\left(\sigma_k^{2} + \sigma^{2} \right)} \right)
\end{equation}
the log-likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$. In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \quad \text{and} \quad \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.
